---
title: "Analysis of Age Distribution of Platform Workers"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    theme: united
---

### Introduction

This notebook analyzes the age distribution of platform workers based on the `trabalhadoresAPP_2020` dataset. A histogram will be generated to visualize the frequency of ages and a boxplot for a statistical summary.

### 1. Load Libraries

First, we load the necessary packages for data analysis and visualization.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # Contains ggplot2 for charts and dplyr for manipulation
library(gridExtra) # Useful for arranging multiple charts in a single view
```

### 2. Data Preparation

In this step, the `trabalhadoresAPP_2020` dataframe is prepared for analysis. We remove any `NA` values in the `Idade` column and ensure it is a numeric type. Then, we calculate the average age of the sample.

```{r data_preparation}
# This code assumes that 'trabalhadoresAPP_2020' already exists in the environment.

# Filter data, ensure the 'Idade' column is numeric
idade_data <- trabalhadoresAPP_2020 %>%
  filter(!is.na(Idade)) %>%
  mutate(Idade = as.numeric(Idade))

# Calculate the average age for use in the charts
idade_media <- mean(idade_data$Idade, na.rm = TRUE)

# Display the calculated average age
print(paste("The average age of the sample is:", round(idade_media, 2), "years."))
```

### 3. Chart 1: Histogram of Age Distribution

The histogram below shows the number of workers in each age group. The visualization has been enhanced according to the BBC R Cookbook's best practices for clarity and impact. A highlighted band has been added to emphasize the primary age group between 30 and 40 years old.

```{r histogram_plot, fig.width=10, fig.height=6}
histograma_idade <- ggplot(idade_data, aes(x = Idade)) +
  # Add a rectangle to highlight the 30-40 age band
  geom_rect(
    aes(xmin = 30, xmax = 40, ymin = -Inf, ymax = Inf),
    fill = "#e76f51", alpha = 0.02 # Using a low alpha to make it a subtle background highlight
  ) +
  geom_histogram(binwidth = 1, fill = "#1380A1", color = "white", alpha = 0.9) +
  geom_vline(
    aes(xintercept = idade_media),
    color = "#e76f51",
    linetype = "dashed",
    size = 1
  ) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") + # Add a base line for the X-axis
  annotate(
    "text",
    x = idade_media + 10,
    y = 700,
    label = paste("Average Age:", round(idade_media, 1), "years"),
    color = "#e76f51",
    size = 4.5,
    fontface = "bold"
  ) +
  # Annotation for the highlighted band
  annotate(
    "text",
    x = 52,
    y = 750,
    label = "Main age group: 30-40 years",
    color = "#e76f51",
    size = 4.5,
    fontface = "bold"
  ) +
  labs(
    title = "Distribution of Platform-based workers by Age",
    subtitle = "Most of the workers are between 30 and 40 years old.",
    x = "Age (years)",
    y = "Number of Workers",
    caption = "Source: PNAD-COVID19 Data (IBGE), 2020."
  ) +
  # Add scale for x-axis to control the labels
  scale_x_continuous(breaks = seq(0, 90, by = 10)) +
  # Theme inspired by the BBC R Cookbook for a clean and professional look
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0, size = 20, face = "bold", color = "#222222"),
    plot.subtitle = element_text(hjust = 0, size = 16, margin = margin(t = 5, b = 20), color = "grey40"),
    plot.caption = element_text(hjust = 1, size = 12, color = "grey40", face = "italic"),
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(size = 12, color = "#333333"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(margin = margin(t = 10))
  )

# Display the chart
print(histograma_idade)
```

### 4. Chart 2: Box Plot for Statistical Summary

The box plot provides a statistical summary of the age distribution, showing the median, quartiles, and outliers.

```{r boxplot_plot, fig.width=8, fig.height=6}
boxplot_idade <- ggplot(idade_data, aes(x = "", y = Idade)) +
  geom_boxplot(fill = "#009E73", alpha = 0.7, width = 0.5) +
  geom_jitter(color = "black", alpha = 0.1, width = 0.15) +
  labs(
    title = "Statistical Summary of Age",
    subtitle = "Median, quartiles, and outliers",
    x = "",
    y = "Age (years)"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Display the chart
print(boxplot_idade)
```

### 5. Combined Visualization (Optional)

For a direct comparison, the two charts can be displayed side-by-side.

```{r combined_plot, fig.width=12, fig.height=6}
# grid.arrange(histograma_idade, boxplot_idade, ncol = 2)
