---
title: "Analysis of Education Level of Platform Workers"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    theme: united
---

### Introduction

This notebook analyzes the education level of platform workers based on the `trabalhadoresAPP_2020` dataset. A waffle chart will be generated to visualize the composition of the workforce across different educational backgrounds.

### 1. Load Libraries

First, we load the necessary packages for data analysis and visualization. `waffle` is added to create the waffle chart.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # Contains ggplot2 for charts and dplyr for manipulation
library(RColorBrewer) # For color palettes
library(waffle)       # For creating waffle charts
```

### 2. Data Preparation

In this step, the `trabalhadoresAPP_2020` dataframe is prepared for analysis. The education levels are translated to English. For the waffle chart, we need a named vector of integer values. The names of the vector are formatted to include the percentage, which will be displayed in the legend.

```{r data_preparation_education}
# This code assumes that 'trabalhadoresAPP_2020' already exists in the environment.

# Check if the 'Escolaridade' column exists
if (!"Escolaridade" %in% names(trabalhadoresAPP_2020)) {
  stop("ERROR: The 'Escolaridade' column was not found in the dataframe.")
}

# Process and translate education data for the waffle chart
education_data <- trabalhadoresAPP_2020 %>%
  filter(!is.na(Escolaridade) & Escolaridade != "Ignorado") %>%
  # Translate education levels to English
  mutate(
    Education_Level = case_when(
      Escolaridade == "Médio completo" ~ "High school complete",
      Escolaridade == "Fundamental incompleto" ~ "Incomplete middle school",
      Escolaridade == "Fundamental completa" ~ "Middle school complete",
      Escolaridade == "Médio incompleto" ~ "Incomplete high school",
      Escolaridade == "Superior completo" ~ "Higher education complete",
      Escolaridade == "Superior incompleto" ~ "Incomplete higher education",
      Escolaridade == "Sem instrução" ~ "Illiterate",
      Escolaridade == "Pós-graduação, mestrado ou doutorado" ~ "Post-graduate degree",
      TRUE ~ as.character(Escolaridade) # Fallback for any other categories
    )
  ) %>%
  count(Education_Level, name = "count") %>%
  # Arrange from largest to smallest
  arrange(desc(count))

# Create a named vector where values are percentages rounded to the nearest integer
total_count <- sum(education_data$count)
waffle_vector <- round((education_data$count / total_count) * 100)

# Due to rounding, the sum might not be exactly 100. We adjust the largest category to fix this.
diff <- 100 - sum(waffle_vector)
waffle_vector[1] <- waffle_vector[1] + diff

# Create new names for the legend that include the percentage values
names(waffle_vector) <- paste(education_data$Education_Level, paste0("(", waffle_vector, "%)"))

# Display the prepared vector
print("Data vector for waffle chart (values are percentages and should sum to 100):")
print(waffle_vector)
```

### 3. Chart: Waffle Chart of Education Levels

The waffle chart below visualizes the workforce composition, where each square represents 1% of the workers. This provides an intuitive way to see the proportion of each education level. The chart is oriented horizontally for readability.

```{r barchart_plot, fig.width=12, fig.height=8}
# Create the waffle chart
# The waffle function is part of the 'waffle' package.
education_chart <- waffle(
  waffle_vector,
  rows = 5,           # Reduced number of rows for a horizontal layout
  size = 2,         # The size of the squares
  flip = FALSE,        # Flips the chart to be horizontal
  colors = RColorBrewer::brewer.pal(length(waffle_vector), "Paired"), # Color palette
  legend_pos = "top" # Position of the legend
) +
# Add titles and annotations using ggplot2's labs function
labs(
  title = "Workforce Composition by Education Level",
  subtitle = "62% of workers have, at most, a complete high school education.",
  x = "1 square = 1% of the workforce",
  caption = "Source: PNAD-COVID19 Data (IBGE), 2020."
) +
theme(
  plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
  plot.subtitle = element_text(hjust = 0.5, size = 16, margin = margin(b = 20)),
  plot.caption = element_text(hjust = 1, face = "italic"),
  legend.text = element_text(size = 11)
)


# Display the chart
print(education_chart)
```