---
title: "Analysis of INSS Contribution of Platform Workers"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    theme: united
---

### Introduction

This notebook analyzes the social security (INSS) contribution status of platform workers based on the `trabalhadoresAPP_2020` dataset. A bar chart will be generated to visualize the proportion of workers who contribute versus those who do not.

### 1. Load Libraries

First, we load the necessary packages for data analysis and visualization.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # Contains ggplot2 for charts and dplyr for manipulation
```

### 2. Data Preparation

In this step, the `trabalhadoresAPP_2020` dataframe is prepared. We will use the `Contribui para o INSS` column, translate its values to English, count the number of workers for each category, and calculate their percentage share.

```{r data_preparation_inss}
# This code assumes that 'trabalhadoresAPP_2020' already exists in the environment.

# Check if the column exists
if (!"Contribui para o INSS" %in% names(trabalhadoresAPP_2020)) {
  stop("ERROR: The 'Contribui para o INSS' column was not found in the dataframe.")
}

# Process the INSS contribution data
inss_data <- trabalhadoresAPP_2020 %>%
  # Filter out missing values, which are coded as "Não aplicável"
  filter(!is.na(`Contribui para o INSS`) & `Contribui para o INSS` != "Não aplicável") %>%
  # Translate the categories to English
  mutate(
    Contribution_Status = case_when(
      `Contribui para o INSS` == "Sim" ~ "Yes",
      `Contribui para o INSS` == "Não" ~ "No"
    )
  ) %>%
  # Count occurrences of each English category
  count(Contribution_Status, name = "count") %>%
  # Calculate the percentage for each level
  mutate(
    percentage = (count / sum(count)) * 100,
    # Reorder the factor levels for a logical plot
    Contribution_Status = fct_reorder(Contribution_Status, count)
  )

# Display the prepared data
print("INSS contribution data ready for analysis:")
print(inss_data)
```

### 3. Chart: Bar Chart of INSS Contribution

The bar chart below shows the number of workers by their INSS contribution status. This visualization highlights the significant proportion of platform workers who do not contribute to the social security system.

```{r barchart_plot_inss, fig.width=10, fig.height=7}
inss_chart <- ggplot(inss_data, aes(x = Contribution_Status, y = count)) +
  # Use geom_col for pre-counted data
  geom_col(fill = "#1380A1", alpha = 0.9, width = 0.5) +
  # Add text labels for the percentage on top of each bar
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
    vjust = -0.5, # Position text above the bar
    size = 5,
    color = "#333333",
    fontface = "bold"
  ) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  # Adjust the y-axis limit to make space for the labels
  scale_y_continuous(limits = c(0, max(inss_data$count) * 1.1)) +
  labs(
    title = "Distribution of workers by social security contribution status",
    subtitle = "Vast Majority of Platform Workers Do Not Contribute to INSS",
    x = "Contributes to INSS?",
    y = "Number of Workers",
    caption = "Source: PNAD-COVID19 Data (IBGE), 2020"
  ) +
  # Theme inspired by the BBC R Cookbook
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0, size = 20, face = "bold", color = "#222222"),
    plot.subtitle = element_text(hjust = 0, size = 16, margin = margin(t = 5, b = 20), color = "grey40"),
    plot.caption = element_text(hjust = 1, size = 12, color = "grey40", face = "italic"),
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(size = 12, color = "#333333"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(margin = margin(t = 10))
  )

# Display the chart
print(inss_chart)
```